Calendar-based graphics for visualising people's daily schedules

### 2017-07-21 main.Rmd
spelling
added the last figure to have chinese labels

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualising people's daily schedules"
  plain: "Calendar-based graphics for visualising people's daily schedules"
abstract: >
  This paper describes a \code{frame\_calendar} function that organises and displays temporal data, collected on sub-daily resolution, into a calendar layout. Calendars are broadly used in society to display temporal information, and events. The \code{frame\_calendar} utilizes linear algebra on the date  variable to create the layout. It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronises neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior, based on counts which are captured at hourly interval by sensors scattered around the city. Facetting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to  events such as work days, weekends, holidays, and special events. The layout algorithm has several format variations and options. It is implemented in the R package \pkg{sugrrants}.

keywords:
  # at least one keyword must be supplied
  formatted: [data visualisation, statistical graphics, time series, R package, grammar of graphics]
  plain:     [data visualisation, statistical graphics, time series, R package, grammar of graphics]
preamble: >
  \usepackage{amsmath}
output: 
  rticles::jss_article:
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

This work was originally motivated by studying foot traffic in the city of Melbourne [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area up until the end of 2016 (see Figure \ref{fig:ped-map}). The dataset can shed light into understanding people's daily schedules, or assisting administration and business planning. We start off with the conventional time series plot to catch a glimpse of such data. A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of the foot traffic at 3 different sensors. Figure \ref{fig:facet-time} provides more details of the temporal patterns by facetting on day of the week. The sensor data, like many temporal datasets on human behaviours, lends itself to a number of exploratory data visualisation challenges:

@@

There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calander plot:

@@

The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}. For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street station has higher pedesterian counts during the weekends and public holidays. This suggests that Flagstaff station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting. Figure \ref{fig:overlay} also shows that State Library follows a similar temporal trend as Flinders Street Station on non-work days. The nighttime events, such as White Night and New Year's Eve has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library.

@@

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of facetting calendar plots. This allows comparing the overall structure between sensors, while emphasising indivdual sensor variation.

@@

```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout. Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consitent on non-work days than work days."}
```

@@

>>>>

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualising people's daily schedules"
  plain: "Calendar-based graphics for visualising people's daily schedules"
abstract: >
  This paper describes a \code{frame\_calendar} function that organises and displays temporal data, collected on sub-daily resolution, into a calendar layout. Calendars are broadly used in society to display temporal information, and events. The \code{frame\_calendar} utilizes linear algebra on the date  variable to create the layout. It utilizes the grammar of graphics to create the plots inside each cell, and thus synchronises neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior, based on counts which are captured at hourly interval by sensors scattered around the city. Facetting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to  events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the R package \pkg{sugrrants}.

keywords:
  # at least one keyword must be supplied
  formatted: [data visualisation, statistical graphics, time series, R package, grammar of graphics]
  plain:     [data visualisation, statistical graphics, time series, R package, grammar of graphics]
preamble: >
  \usepackage{amsmath}
  \usepackage{tikz}
output: 
  rticles::jss_article:
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

This work was originally motivated by studying foot traffic in the city of Melbourne [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area up until the end of 2016 (see Figure \ref{fig:ped-map}). The dataset can shed light into understanding people's daily schedules, or assisting administration and business planning. We start off with the conventional time series plot to catch a glimpse of such data. A small multiples, shown in Figure \ref{fig:time-series-plot}, gives an overall picture of the foot traffic at 3 different sensors. Figure \ref{fig:facet-time} provides more details of the temporal patterns by faceting on day of the week. The sensor data, like many temporal datasets on human behaviours, lends itself to a number of exploratory data visualisation challenges:

@@

There are several options provided for the \code{frame_calendar} function to initialize and adjust the display of a calender plot:

@@

The comparison of one sensor to another adds additional insights to the dataset, which is commonly done with an overlaying plot such as Figure \ref{fig:overlay}. For instance, the dominant patterns that occurred at Flinders Street and Flagstaff train stations are driven by the commuters on work days; however, Flinders Street station has higher pedestrian counts during the weekends and public holidays. This suggests that Flagstaff station limits its functionality on non-work days, but various activities take place around the Flinders Street Station other than commuting. Figure \ref{fig:overlay} also shows that State Library follows a similar temporal trend as Flinders Street Station on non-work days. The nighttime events, such as White Night and New Year's Eve has barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and State Library.

@@

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of faceting calendar plots. This allows comparing the overall structure between sensors, while emphasising individual sensor variation.

@@

```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout. Previous hour's count is plotted against each hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days."}
```

@@

```{r chn, dev.args = list(family = "GB1"), fig.height = 6.5, fig.cap = "Figure 12 with Chinese text labels."}
```

### 2017-07-21 main.tex
added the last figure to have chinese labels

\usepackage{amsmath}

@@

This work was originally motivated by studying foot traffic in the city
of Melbourne \citep{ped}. There have been 43 sensors installed that
count pedestrians every hour across the downtown area up until the end
of 2016 (see Figure \ref{fig:ped-map}). The dataset can shed light into
understanding people's daily schedules, or assisting administration and
business planning. We start off with the conventional time series plot
to catch a glimpse of such data. A small multiples, shown in Figure
\ref{fig:time-series-plot}, gives an overall picture of the foot traffic
at 3 different sensors. Figure \ref{fig:facet-time} provides more
details of the temporal patterns by facetting on day of the week. The
sensor data, like many temporal datasets on human behaviours, lends
itself to a number of exploratory data visualisation challenges:

@@

>>>>

\usepackage{amsmath} \usepackage{tikz}

@@

This work was originally motivated by studying foot traffic in the city
of Melbourne \citep{ped}. There have been 43 sensors installed that
count pedestrians every hour across the downtown area up until the end
of 2016 (see Figure \ref{fig:ped-map}). The dataset can shed light into
understanding people's daily schedules, or assisting administration and
business planning. We start off with the conventional time series plot
to catch a glimpse of such data. A small multiples, shown in Figure
\ref{fig:time-series-plot}, gives an overall picture of the foot traffic
at 3 different sensors. Figure \ref{fig:facet-time} provides more
details of the temporal patterns by faceting on day of the week. The
sensor data, like many temporal datasets on human behaviours, lends
itself to a number of exploratory data visualisation challenges:

@@

{\centering \includegraphics[width=\textwidth]{figure/chn-1}

\caption[Figure 12 with Chinese text labels]{Figure 12 with Chinese text labels.}\label{fig:chn}
\end{figure}
\end{CodeChunk}
